Stack Safety is a Security Property

### 2021-04-16 tex/main.tex
Numbered subsections in case study
Moved sections around

%% Journal information
%% Supplied to authors by publisher for camera-ready submission;
%% use defaults for review submission.
\acmJournal{PACMPL}
\acmVolume{1}
\acmNumber{OOPSLA} % CONF = POPL or ICFP or OOPSLA
\acmArticle{1}
\acmYear{2021}
\acmMonth{1}
\acmDOI{} % \acmDOI{10.1145/nnnnnnn.nnnnnnn}
\startPage{1}

@@

To demonstrate the utility of our formal characterization, we use these
properties to validate and improve an existing enforcement mechanism, the
{\em stack-safety micro-policies} of~\citet{DBLP:conf/sp/RoesslerD18}.  We
use QuickChick~\citep{Denes:VSL2014,Pierce:SF4}, a property-based testing
tool for the Coq proof assistant, to generate many random programs and check
that Roessler and Dehon's micro-policies correctly abort the ones that
attempt to violate one of our properties; furthermore, we introduce several
intentionally broken micro-policies and check that the testing framework is
able to generate counterexamples that violate our properties but are not
halted by the incorrect enforcement variants.  (Showing that the testing
framework detects intentional bugs in enforcement mechanisms increases our
confidence that it can also catch unknown ones, strengthening our confidence
in the enforcement mechanism itself.\ifaftersubmission\bcp{Maybe that
  sentence belongs in the testing section?}\fi)
%
We find that Roessler and Dehon's {\em Depth Isolation} micro-policy, in
which memory cells within each stack frame are tagged with the identity of
the function activation that owns the frame and access to those locations is
then permitted only when that activation is currently executing, validates our
stepwise properties.  On the other hand, their more efficient \emph{Lazy
  Tagging and Clearing} policy enforces observational integrity, but testing
reveals\bcp{right, Leo?} that it does not enforce confidentiality (either
stepwise or
observational): it violates the temporal aspect of confidentiality in
corner cases where data can leak across repeated calls to the same callee.  We
propose a variant of {\em Lazy Tagging and Clearing} that \emph{does} enforce
confidentiality (as far as our testing can tell), albeit at some performance
cost.
%
\ifaftersubmission
\bcp{Did Nick and Andre know about this deficiency?  Did they mention
  it in their paper?  (If neither, we should make a bigger deal out of the
  fact that we discovered it.)\apt{Their paper is a little vague on this subject.
    They refer to ``temporal security properties'' as something to do with dangling
    pointers, but they don't define it.
    Then they say that Depth Isolation has weaker temporal security
    guarantees than Static Authorities (tag=function name), but they don't explain why
    (and I don't see why). They don't say anything explicit about the temporal
    security properties of Lazy Tagging and Clearing, but they consider it to enforce
    a completely different (and generally weaker) policy anyway (``data flow
    integrity'' vs. ``memory safety'').}}\bcp{Sigh.  OK, I propose that we
  leave this issue to deal with later (in consultation with them, which we
  don't have time to do a good job of today).  At any rate, by my reading we
are not saying anything incorrect at the moment.}
\fi

@@

\begin{itemize}
\item We adapt concepts of integrity and confidentiality from language-based
security, combine these with an existing notion of well-bracketed
control flow\bcp{need to mention this above} to formalize a {\em stepwise stack-safety}
property, and argue that this property captures the intuitive concept of
strict stack
safety~(\cref{sec:prelim,sec:lse}).
\item We use property-based randomized testing to check that this stepwise
property is actually enforced by the {\em Depth Isolation} policy
of~\citet{DBLP:conf/sp/RoesslerD18} (\cref{sec:enforcement,sec:testing}).
\bcp{mention that we extend the ICFP gen-by-exec with nested stuff...}
\item We formalize a
weaker but more extensional \emph{observational} variant of the stack safety
property that characterizes a refinement of Roessler and DeHon's more
performant {\em Lazy Tagging and  Clearing} policy (\cref{sec:lazy}).
\item We show that the definition can be
extended to argument passing on the stack and to a simple coroutine system.
\bcp{fill in pointer to new section}
\end{itemize}

@@

\begin{subfigure}[t]{.2\textwidth}
{\small
\begin{verbatim}
int main() {
  int x = 42;
  return f()+x;
}

int f() {
  int y;
  *(&y-2) = 0;
  return 5;
}
\end{verbatim}
}
\end{subfigure}
\begin{subfigure}[t]{.5\textwidth}
{\small
\begin{verbatim}
1 main: add $1,%sp      ; allocate frame
2       mov $42,-1(%sp) ; initialize local
3       call f
4       add -1(%sp),%ra ; set return value
5       sub $1,%sp      ; deallocate frame
6       ret

10 f:   add $1,%sp      ; allocate frame
11      mov $0,-3(%sp)  ; violation!
12      mov $5, %ra     ; set return value
13      sub $1,%sp      ; deallocate frame
14      ret
\end{verbatim}
}
\end{subfigure}
\begin{subfigure}[t]{.25\textwidth}
\begin{center}
\begin{tabular}{l l}
{\tt 2} &
\memoryaddrs{8em}
\memory{3}{\unsealc}[Unseal]
~$\cdots$
\vspace{.5em}
\\
{\tt 3} &
\memoryaddrs{8em}
\memory{3}{\unsealc}
~$\cdots$
    \MemoryLabel{-15em}{0.75em}{42}
    \vspace{.5em}
\\
{\tt 11} &
\memoryaddrs{16em}
\memory{1}{\mainsealc}[Seal(0)]%
\memory{1}{\retptrc}[RetPtr]%
\memory{1}{\unsealc}[Unseal]
~$\cdots$
\MemoryLabel{-15em}{0.75em}{42}
\vspace{.5em}
\\
{\tt 12} &
\memoryaddrs{16em}
\memory{1}{\mainsealc}
\memory{1}{\retptrc}
\memory{1}{\unsealc}
~$\cdots$
\MemoryLabel{-15em}{0.75em}{\bf 0}
\vspace{.5em}
\end{tabular}
\end{center}

\vspace{\abovedisplayskip}
\end{subfigure}
\caption{An integrity violation
\ifaftersubmission\bcp{I wonder if we
    could make the figures easier to read by putting very light, differently
  colored backgrounds under the different parts}\fi}
\label{fig:int1}
\end{figure}

This section introduces the intuition behind the properties through a series
of examples. Existing stack safety enforcement work mostly refers directly to examples
of bad behavior that it prevents; we will argue that: \bcp{Not sure this
  enumeration is useful.}\apt{+1}

@@

\begin{figure}
\begin{subfigure}[t]{.2\textwidth}
{\small
\begin{verbatim}
int main() {
  int x = 42;
  return f()+x;
}

int f() {
  int y,z;
  y = 5;
  z = *(&y-2);
  return y+z;
}
\end{verbatim}
}
\end{subfigure}
\begin{subfigure}[t]{.5\textwidth}
{\small
\begin{verbatim}
1 main: add $1,%sp      ; allocate frame
2       mov $42,-1(%sp) ; initialize local
3       call f
4       add -1(%sp),%ra ; set return value
5       sub $1,%sp      ; deallocate frame
6       ret

@@

\paragraph*{Control-Flow Attacks}
\begin{figure}
  \centering
  \begin{subfigure}[t]{.2\textwidth}
{\footnotesize
\begin{verbatim}
int *stash = 0;
int main() {
  int x = 1;
  f();
  x = -x;
  f();
  return x;
}

void f() {
  int y;
  if (!stash) {
    stash = *(&y-1);
  } else {
    *(&y-1) = stash;
    stash = 0;
  }
  return;
}
\end{verbatim}
}
\end{subfigure}
  \begin{subfigure}[t]{.4\textwidth}
{\footnotesize
\begin{verbatim}
   stash: .word 0
1  main:  add $1,%sp     ; alloc frame
2         mov $1,-1(%sp) ; init x
3         call f
4         neg -1(%sp)    ; x = -x
5         call f
6         mov -1(%sp),%ra ; set ret val
7         sub $1,%sp     ; dealloc frame
6         ret

10 f:     add $1,%sp     ; alloc frame
11        cmp stash, $0  ; stash = 0?
12        breq #15       ; branch if so
13        mov -2(%sp),stash ; set stash
14        jmp #17
15        mov stash,-2(%sp) ; get stash
16        mov $0,stash
17        sub $1,%sp     ; dealloc frame
18        ret
\end{verbatim}
}
  \end{subfigure}
  \begin{subfigure}[t]{.25\textwidth}
    \begin{tabular}{l l l}
      {\tt 3} &
      \memoryaddrs{8em}
      \memory{3}{\unsealc}
      ~$\cdots$
      \MemoryLabel{-15em}{0.75em}{1} \\
      {\tt 18} &
      \memoryaddrs{12em}
      \memory{1}{\mainsealc}[Seal(0)]%
      \memory{1}{\unsealc}[RetPtr]%
      \memory{1}{\unsealc}%
      ~$\cdots$
      \MemoryLabel{-15em}{0.75em}{1}
      \MemoryLabel{-11em}{0.75em}{\#4}
      \vspace{.5em} &
      \hspace*{-1.5em}
      \memory[1.2em]{1}{\mainsealc}[Target]%
      \MemoryLabel{-2.2em}{0.75em}{\#4}
      \\
      {\tt 4} &
      \memoryaddrs{8em}
      \memory{3}{\unsealc}
      ~$\cdots$
      \MemoryLabel{-15em}{0.75em}{1} \\
      {\tt 11} &
      \memoryaddrs{16em}
      \memory{1}{\mainsealc}
      \memory{2}{\unsealc}%
      ~$\cdots$
      \MemoryLabel{-15em}{0.75em}{-1}
      \MemoryLabel{-11em}{0.75em}{\#6} &
      \hspace*{-1em}
      \memory[1.2em]{1}{\mainsealc}
      \MemoryLabel{-1.2em}{0.75em}{\#6}
      \\
      {\tt 18} &
      \memoryaddrs{12em}
      \memory{1}{\mainsealc}
      \memory{1}{\badc}
      \memory{1}{\unsealc}%
      ~$\cdots$
      \MemoryLabel{-15em}{0.75em}{-1}
      \MemoryLabel{-11em}{0.75em}{\#4} &
      \hspace*{-1em}
      \memory[1.2em]{1}{\mainsealc}
      \MemoryLabel{-1.2em}{0.75em}{\#6}
      \\
      {\tt 4} &
      \memoryaddrs{8em}
      \memory{1}{\mainsealc}
      \memory{2}{\unsealc}
      ~$\cdots$
      \MemoryLabel{-15em}{0.75em}{-1} &
      \hspace*{-1em}
      \memory[1.2em]{1}{\mainsealc}
      \MemoryLabel{-1.2em}{0.75em}{\#6}
      \\
      {\tt 5} &
      \memoryaddrs{8em}
      \memory{1}{\mainsealc}
      \memory{2}{\unsealc}
      ~$\cdots$
      \MemoryLabel{-15em}{0.75em}{\bf 1} &
      \hspace*{-1em}
      \memory[1.2em]{1}{\mainsealc}
      \MemoryLabel{-1.2em}{0.75em}{\#6} \\
    \end{tabular}
  \end{subfigure}
  \caption{A control-flow attack}
  \label{fig:controlflow}
\end{figure}

@@

Together, integrity and confidentiality protect a caller's data until its callee returns
to it. However, it is possible that an intermediate caller is not returned
{\em to}, but rather returned {\em past}, to some ancestor caller! For
example, function $A$ calls function B, which in turn calls $C$, which
returns directly to $A$ by resetting the stack pointer to $A$'s stack frame
and branching to the address after the call from $A$ to $B$.
Strictly speaking, this is not an information-flow problem: the caller's XXXXXX.
\BCP{Still working on this bit.}
But a lower level caller might have been relying on the intermediate
caller to finish its computation, so it is worth protecting control flow even from
attacks that do not violate data protections. We formalize this as {\em return-to-caller}.

@@

\begin{figure}
  \centering
  \begin{subfigure}[t]{.2\textwidth}
{\small
\begin{verbatim}
void main() {
  int x = 5;
  x = f(x);
  return;
}

void f(int a) {
  int y = 2*a;
  a = 0;
  return;
}
\end{verbatim}
}
  \end{subfigure}
  \begin{subfigure}[t]{.6\textwidth}
{\small
\begin{verbatim}
1  main: add $1,%sp          ; allocate frame
2        mov $5,-1(%sp)      ; initalize x
3        add $1,%sp          ; allocate arg
4        mov -2(%sp),-1(%sp) ; set arg
5        call f              ; SHARE: %sp-1
6        sub $1,%sp          ; deallocate arg
7        mov %ra,-1(%sp)     ; store x
8        sub $1,%sp          ; deallocate frame
9        ret

10  f:   add $1,%sp          ; allocate frame
11       mov -3(%sp),-1(%sp) ; y = a
12       mul $2,-1(%sp)      ; y = a*2
13       mov $0,-3(%sp)      ; a = 0
14       sub $1,%sp          ; deallocate frame
15       ret
\end{verbatim}
}
  \end{subfigure}
\begin{subfigure}{.65\textwidth}
\vspace*{0.2in}
\begin{center}
\begin{tabular}{l r | l}
  {\tt 3} &
  \multicolumn{2}{c}{
    \memoryaddrs[(0)]{8em}
    \memory{4}{\unsealc}
    ~$\cdots$
    \MemoryLabel{-19em}{0.75em}{5}
%    \MemoryLabel{-14em}{0.75em}{0}
%    \MemoryLabel{-10em}{0.75em}{0}
%    \MemoryLabel{-7em}{0.75em}{0}
    \vspace{.5em}
  } \\
  {\tt 11} &
  \memoryaddrs[(1)]{21.5em}
  \memory{1}{\mainsealc}[Seal(0)]%
  \memory{1}{\mainpassc}[P(0,1)]%
  \memory{1}{\retptrc}[RetPtr]%
  \memory{1}{\unsealc}
  ~$\cdots$
  \MemoryLabel{-19em}{0.75em}{5}
  \MemoryLabel{-14em}{0.75em}{5}
%  \MemoryLabel{-6em}{0.75em}{0}
  &
  \memoryaddrs[(1)]{21.5em}
  \memory{1}{\mainsealc}
  \memory{1}{\mainpassc}
  \memory{1}{\retptrc}
  \memory{1}{\unsealc}
  ~$\cdots$
  \MemoryLabel{-19em}{0.75em}{\(v_0\)}
  \MemoryLabel{-14em}{0.75em}{5}
  \MemoryLabel{-6em}{0.75em}{\(v_1\)}
  \\
  {\tt 13} &
  \memoryaddrs[(1)]{21.5em}
  \memory{1}{\mainsealc}
  \memory{1}{\mainpassc}
  \memory{1}{\retptrc}
  \memory{1}{\unsealc}
  ~$\cdots$
  \MemoryLabel{-19em}{0.75em}{5}
  \MemoryLabel{-14em}{0.75em}{5}
  \MemoryLabel{-6em}{0.75em}{10}
  &
  \memoryaddrs[(1)]{21.5em}
  \memory{1}{\mainsealc}
  \memory{1}{\mainpassc}
  \memory{1}{\retptrc}
  \memory{1}{\unsealc}
  ~$\cdots$
  \MemoryLabel{-19em}{0.75em}{\(v_0\)}
  \MemoryLabel{-14em}{0.75em}{5}
  \MemoryLabel{-6em}{0.75em}{10}
  \\
  {\tt 14} &
  \memoryaddrs[(1)]{21.5em}
  \memory{1}{\mainsealc}
  \memory{1}{\mainpassc}
  \memory{1}{\retptrc}
  \memory{1}{\unsealc}
  ~$\cdots$
  \MemoryLabel{-19em}{0.75em}{5}
  \MemoryLabel{-14em}{0.75em}{0}
  \MemoryLabel{-6em}{0.75em}{10}
  &
  \memoryaddrs[(1)]{21.5em}
  \memory{1}{\mainsealc}
  \memory{1}{\mainpassc}
  \memory{1}{\retptrc}
  \memory{1}{\unsealc}
  ~$\cdots$
  \MemoryLabel{-19em}{0.75em}{\(v_0\)}
  \MemoryLabel{-14em}{0.75em}{0}
  \MemoryLabel{-6em}{0.75em}{10}
  \\
  \\
  {\tt 8} &
  \multicolumn{2}{c}{
    \memoryaddrs[(0)]{8em}
    \memory{4}{\unsealc}
    ~$\cdots$
    \MemoryLabel{-19em}{0.75em}{10}
    \MemoryLabel{-14em}{0.75em}{0}
    \MemoryLabel{-10em}{0.75em}{\#6}
    \MemoryLabel{-7em}{0.75em}{10}
    \vspace{.5em}
  } \\
\end{tabular}
\end{center}

To this point we have ignored function arguments, since our simplest model places them in
registers. This is an oversimplification, which we will now correct, aiming at two kinds
of sharing: passing by value on the stack, and passing by reference. First we will explore
some examples at the same level of detail as our earlier ones, and then move on to
extending our formalization.

@@

In this section we explore how an existing enforcement mechanism, {\em
  Depth Isolation}, from \citet{DBLP:conf/sp/RoesslerD18} implements
the formal stack-safety property described in the previous
section. (In fact, it enforces a slightly stronger, inductive version
of that property which strengthens the local state encapsulation
component by checking for integrity and confidentiality violations at
every step of the callee's execution trace.) We validate this claim
with a property-based testing framework that can quickly uncover
inconsistencies between the enforcement mechanism and the stack-safety
property.

@@

The micro-policy we present here to enforce the definition of stack
safety developed in \cref{sec:lse} is a small variation
on the {\em Depth Isolation} policy presented by
\citet{DBLP:conf/sp/RoesslerD18}. In their policy,  the stack memory is tagged
with ownership information associated to each stack frame (and to each
separate object inside that frame, a more finely grained access
control that we do not need to consider here), and tags registers containing
stack pointers with access permission information. The policy also
uses tags on instructions to identify the code
sequences that have permission to manage the stack, say during calls and returns.
\leo{If we have the simple exposition so far, no need to mention the following sentence:}
In the following, for ease of exposition, we will assume that no
arguments are passed on the stack.
%% \rb{Conceivably on alloc
%%   operations, etc., which we have not yet mentioned.}

@@

\item From callee back to caller, when execution is about to reach a
  return point.
  %The exit sequence retags all locations in the function frame with $\tagNoDepth$.
  At the actual return instruction,
  the $\PCname$ tag is decremented.
%  \leo{Not sure the actual policy
%    does that now (retag unused). But I don't think it's a
%    confidentiality or integrity violation.}
%  \rb{Correct, and it doesn't. Some versions of the policy do clear
%    the stack, retagging would be easily added.}

@@

We validated the micro-policy of the previous section by
systematically testing that it satisfies eager stack integrity and
confidentiality. We used a Coq specification of the RISC-V
architecture
\footnote{Available at \url{https://github.com/mit-plv/riscv-coq}.}
\leo{translated from Haskell via hs-2-coq?}  and extended it with a
runtime monitor implementing the stack-safety
micro-policy.\footnote{This code will be made public. \leo{More?}} We
  chose the Coq proof assistant as the setting for our implementation
  for two reasons: first, to ensure that our coinductive trace
  definitions are well-formed, and reason about them\leo{Do we prove
    anything interesting anymore?}, and to leverage the power of the
  QuickChick property-based testing framework~\citep{Pierce:SF4}.

@@

To write such generators we build on the work of
\citep{TestingNI:ICFP, DBLP:journals/jfp/HritcuLSADHPV16}, where they
introduced {\em generation by execution} to produce progams leading
abstract machines towards interesting behaviors. Generation by
execution receives as an input a partially instantiated machine state
and attempts to generate an instruction (or a sequence of instructions
such as a blessed sequence) that makes sense locally (e.g., jumps go
to a potentially valid code location and loads read from from a
potentially valid stack location). Then we step the machine and repeat
the process until we generate or execute some target number of
instructions, or if we reach a point where the machine can't step
anymore.

@@

In addition, we needed to extend this technique to handle the nested
nature of our confidentiality property: rather than generating two
initial machines that are variant of one another and letting them
execute to test for noninterference, we need to generate a new variant
{\em every time a call is made} and check confidentiality for the
subtrace produced from that variant state until its corresponding
return. As a result, a ``single'' confidentiality test compactly
checks the confidentiality of multiple nested calls.

@@

To ensure the effectiveness of testing against our formal properties, we
used {\em mutation testing}~\citep{JiaH11}. In mutation testing, we inject errors
(mutations) in a program that should cause the property of interest (here,
stack safety) to fail, and ensure that the testing framework can find
them. The bugs we used for our evaluation were either artificially generated
by us (deliberately weakening the micro-policy in ways that we expected
should break its guarantees), or actual bugs that we discovered through
testing our implementation. We elaborate on some such bugs below.

\begin{table}[]
\centering
\begin{tabular}{c|c|c}
  Bug & Property Violated & MTTF (ms) \\
  \hline
      {\em LOAD\_NO\_CHECK}  & Confidentiality & 1670 \\
      {\em STORE\_NO\_CHECK} & Integrity & 638 \\
      {\em HEADER\_NO\_INIT} & Integrity & 1488 \\
\end{tabular}
\vspace*{1em}
\caption{MTTF for finding bugs in erroneous policy enforcement mechanisms}
\vspace*{-2em}
\label{tab:bug-table}
\end{table}

>>>>

%% Journal information
%% Supplied to authors by publisher for camera-ready submission;
%% use defaults for review submission.
\acmJournal{PACMPL}
\acmVolume{1}
\acmNumber{SPLASH} % CONF = POPL or ICFP or OOPSLA
\acmArticle{1}
\acmYear{2021}
\acmMonth{1}
\acmDOI{} % \acmDOI{10.1145/nnnnnnn.nnnnnnn}
\startPage{1}

@@

To demonstrate the utility of our formal characterization, we use these
properties to validate and improve an existing enforcement mechanism, the
{\em stack-safety micro-policies} of~\citet{DBLP:conf/sp/RoesslerD18}.  We
use QuickChick~\citep{Denes:VSL2014,Pierce:SF4}, a property-based testing
tool for the Coq proof assistant, to generate many random programs and check
that Roessler and Dehon's micro-policies correctly abort the ones that
attempt to violate one of our properties; furthermore, we introduce several
intentionally broken micro-policies and check that the testing framework is
able to generate counterexamples that violate our properties but are not
halted by the incorrect enforcement variants.  (Showing that the testing
framework detects intentional bugs in enforcement mechanisms increases our
confidence that it can also catch unknown ones, strengthening our confidence
in the enforcement mechanism itself.\ifaftersubmission\bcp{Maybe that
  sentence belongs in the testing section?}\fi)
%
We find that Roessler and Dehon's {\em Depth Isolation} micro-policy, in
which memory cells within each stack frame are tagged with the identity of
the function activation that owns the frame and access to those locations is
then permitted only when that activation is currently executing, validates our
stepwise properties.  On the other hand, their more efficient \emph{Lazy
  Tagging and Clearing} policy enforces observational integrity, but testing
reveals\bcp{right, Leo?} that it does not enforce confidentiality (either
stepwise or
observational): it violates the temporal aspect of confidentiality in
corner cases where data can leak across repeated calls to the same callee.  We
propose a variant of {\em Lazy Tagging and Clearing} that \emph{does} enforce
confidentiality (as far as our testing can tell), albeit at some performance
cost.  \bcp{Did Nick and Andre know about this deficiency?  Did they mention
  it in their paper?  (If neither, we should make a bigger deal out of the
  fact that we discovered it.)\apt{Their paper is a little vague on this subject.
    They refer to ``temporal security properties'' as something to do with dangling
    pointers, but they don't define it.
    Then they say that Depth Isolation has weaker temporal security
    guarantees than Static Authorities (tag=function name), but they don't explain why
    (and I don't see why). They don't say anything explicit about the temporal
    security properties of Lazy Tagging and Clearing, but they consider it to enforce
    a completely different (and generally weaker) policy anyway (``data flow
    integrity'' vs. ``memory safety'').}}\bcp{Sigh.  OK, I propose that we
  leave this issue to deal with later (in consultation with them, which we
  don't have time to do a good job of today).  At any rate, by my reading we
are not saying anything incorrect at the moment.}

@@

\begin{itemize}
\item We adapt concepts of integrity and confidentiality from language-based
security to formalize a {\em stepwise stack-safety}
property and argue that it captures the intuitive concept of strict stack
safety~(\cref{sec:prelim,sec:lse}).
\item We use property-based randomized testing to check that this stepwise
property is actually enforced by the {\em Depth Isolation} policy
of~\citet{DBLP:conf/sp/RoesslerD18} (\cref{sec:enforcement,sec:testing}).
\item We formalize a
weaker but more extensional \emph{observational} variant of the stack safety
property that characterizes a refinement of Roessler and DeHon's more
performant {\em Lazy Tagging and  Clearing} policy (\cref{sec:lazy}).
\item We show that these definitions can be
  extended to argument passing on the stack and to a simple coroutine system
  (\cref{sec:ext}).
\end{itemize}

@@

\begin{subfigure}{.2\textwidth}
{\small
\begin{verbatim}
int main() {
  int x = 42;
  return f()+x;
}

int f() {
  int y;
  *(&y-2) = 0;
  return 5;
}
\end{verbatim}
}
\end{subfigure}
\begin{subfigure}{.5\textwidth}
{\small
\begin{verbatim}
1 main: add $1,%sp      ; allocate frame
2       mov $42,-1(%sp) ; initialize local
3       call f
4       add -1(%sp),%ra ; set return value
5       sub $1,%sp      ; deallocate frame
6       ret

10 f:   add $1,%sp      ; allocate frame
11      mov $0,-3(%sp)  ; violation!
12      mov $5, %ra     ; set return value
13      sub $1,%sp      ; deallocate frame
14      ret
\end{verbatim}
}
\end{subfigure}
\begin{subfigure}{.25\textwidth}
\begin{center}
\begin{tabular}{l l}
{\tt 2} &
\memoryaddrs{8em}
\memory{3}{\unsealc}[Unseal]
~$\cdots$
\vspace{.5em}
\\
{\tt 3} &
\memoryaddrs{8em}
\memory{3}{\unsealc}
~$\cdots$
    \MemoryLabel{-15em}{0.75em}{42}
    \vspace{.5em}
\\
{\tt 11} &
\memoryaddrs{16em}
\memory{1}{\mainsealc}[Seal(0)]%
\memory{1}{\retptrc}[RetPtr]%
\memory{1}{\unsealc}[Unseal]
~$\cdots$
\MemoryLabel{-15em}{0.75em}{42}
\vspace{.5em}
\\
{\tt 12} &
\memoryaddrs{16em}
\memory{1}{\mainsealc}
\memory{1}{\retptrc}
\memory{1}{\unsealc}
~$\cdots$
\MemoryLabel{-15em}{0.75em}{\bf 0}
\vspace{.5em}
\end{tabular}
\end{center}

\vspace{\abovedisplayskip}
\end{subfigure}
\caption{An integrity violation\bcp{Can we align the C-like snippets with
    the corresponding assembly?\apt{Maybe, but it would require adding blank lines
      at places in both the C and asssembly, making them harder to read
  and eating more space.}\bcp{I'm just talking about removing the space
  above "main" in the left column and moving f down a bit.}}
\ifaftersubmission\bcp{I wonder if we
    could make the figures easier to read by putting very light, differently
  colored backgrounds under the different parts}\fi}
\label{fig:int1}
\end{figure}

This section introduces the intuition behind the properties through a series
of examples. Existing stack safety enforcement work mostly refers directly to examples
of bad behavior that it prevents; we will argue that: \bcp{Not sure this
  enumeration is useful.}

@@

\begin{figure}
\begin{subfigure}{.2\textwidth}
{\small
\begin{verbatim}
int main() {
  int x = 42;
  return f()+x;
}

int f() {
  int y,z;
  y = 5;
  z = *(&y-2);
  return y+z;
}
\end{verbatim}
}
\end{subfigure}
\begin{subfigure}{.5\textwidth}
{\small
\begin{verbatim}
1 main: add $1,%sp      ; allocate frame
2       mov $42,-1(%sp) ; initialize local
3       call f
4       add -1(%sp),%ra ; set return value
5       sub $1,%sp      ; deallocate frame
6       ret

@@

\paragraph*{Control-Flow Attacks}
\begin{figure}
  \centering
  \begin{subfigure}{.2\textwidth}
{\footnotesize
\begin{verbatim}
int *stash = 0;

int main() {
  int x = 1;
  f();
  x = -x;
  f();
  return x;
}

void f() {
  int y;
  if (!stash) {
    stash = *(&y-1);
  } else {
    *(&y-1) = stash;
    stash = 0;
  }
  return;
}
\end{verbatim}
}
\end{subfigure}
  \begin{subfigure}{.4\textwidth}
{\footnotesize
\begin{verbatim}
   stash: .word 0
1  main:  add $1,%sp     ; alloc frame
2         mov $1,-1(%sp) ; init x
3         call f
4         neg -1(%sp)    ; x = -x
5         call f
6         mov -1(%sp),%ra ; set ret val
7         sub $1,%sp     ; dealloc frame
6         ret

10 f:     add $1,%sp     ; alloc frame
11        cmp stash, $0  ; stash = 0?
12        breq #15       ; branch if so
13        mov -2(%sp),stash ; set stash
14        jmp #17
15        mov stash,-2(%sp) ; get stash
16        mov $0,stash
17        sub $1,%sp     ; dealloc frame
18        ret
\end{verbatim}
}
  \end{subfigure}
  \begin{subfigure}{.25\textwidth}
    \begin{tabular}{l l l}
      {\tt 3} &
      \memoryaddrs{8em}
      \memory{3}{\unsealc}
      ~$\cdots$
      \MemoryLabel{-15em}{0.75em}{1} \\
      {\tt 18} &
      \memoryaddrs{12em}
      \memory{1}{\mainsealc}[Seal(0)]%
      \memory{1}{\unsealc}[RetPtr]%
      \memory{1}{\unsealc}%
      ~$\cdots$
      \MemoryLabel{-15em}{0.75em}{1}
      \MemoryLabel{-11em}{0.75em}{\#4}
      \vspace{.5em} &
      \hspace*{-1.5em}
      \memory[1.2em]{1}{\mainsealc}[Target]%
      \MemoryLabel{-2.2em}{0.75em}{\#4}
      \\
      {\tt 4} &
      \memoryaddrs{8em}
      \memory{3}{\unsealc}
      ~$\cdots$
      \MemoryLabel{-15em}{0.75em}{1} \\
      {\tt 11} &
      \memoryaddrs{16em}
      \memory{1}{\mainsealc}
      \memory{2}{\unsealc}%
      ~$\cdots$
      \MemoryLabel{-15em}{0.75em}{-1}
      \MemoryLabel{-11em}{0.75em}{\#6} &
      \hspace*{-1em}
      \memory[1.2em]{1}{\mainsealc}
      \MemoryLabel{-1.2em}{0.75em}{\#6}
      \\
      {\tt 18} &
      \memoryaddrs{12em}
      \memory{1}{\mainsealc}
      \memory{1}{\badc}
      \memory{1}{\unsealc}%
      ~$\cdots$
      \MemoryLabel{-15em}{0.75em}{-1}
      \MemoryLabel{-11em}{0.75em}{\#4} &
      \hspace*{-1em}
      \memory[1.2em]{1}{\mainsealc}
      \MemoryLabel{-1.2em}{0.75em}{\#6}
      \\
      {\tt 4} &
      \memoryaddrs{8em}
      \memory{1}{\mainsealc}
      \memory{2}{\unsealc}
      ~$\cdots$
      \MemoryLabel{-15em}{0.75em}{-1} &
      \hspace*{-1em}
      \memory[1.2em]{1}{\mainsealc}
      \MemoryLabel{-1.2em}{0.75em}{\#6}
      \\
      {\tt 5} &
      \memoryaddrs{8em}
      \memory{1}{\mainsealc}
      \memory{2}{\unsealc}
      ~$\cdots$
      \MemoryLabel{-15em}{0.75em}{\bf 1} &
      \hspace*{-1em}
      \memory[1.2em]{1}{\mainsealc}
      \MemoryLabel{-1.2em}{0.75em}{\#6} \\
    \end{tabular}
  \end{subfigure}
  \caption{A control-flow attack}
  \label{fig:controlflow}
\end{figure}

@@

Together, integrity and confidentiality protect a caller's data until its callee returns
to it. But it is possible that an intermediate caller is never returned to, but returned
{\em past}, to a deeper caller. Fundamentally, this is a control-flow problem, not a
data-protection problem. From that caller's perspective, it is no different from the
callee diverging. But a lower level caller might have been relying on the intermediate
caller to finish its computation, so it is worth protecting control flow even from
attacks that do not violate data protections. We formalize this as {\em return-to-caller}.

@@

@@

In this section we sketch an existing enforcement mechanism, {\em Depth
Isolation}, from \citet{DBLP:conf/sp/RoesslerD18}, that implements the
formal stack-safety property described in the previous section. (In
fact, it enforces a slightly stronger, inductive version of that property which
strengthens the local state encapsulation component by checking for
integrity and confidentiality violations at every step of the callee's
execution trace.) We validate this claim with a property-based
testing framework that can quickly uncover inconsistencies between the
enforcement mechanism and the stack-safety property.

@@

The micro-policy we present here to enforce the definition of stack
safety developed in \cref{sec:lse} is a small variation
on the {\em Depth Isolation} policy presented by
\citet{DBLP:conf/sp/RoesslerD18}. In their policy,  the stack memory is tagged
with ownership information associated to each stack frame (and to each
separate object inside that frame, a more finely grained access
control that we do not need to consider here), and tags registers containing
stack pointers with access permission information. The policy also
uses tags on instructions to identify the code
sequences that have permission to manage the stack, say during calls and returns.
In the following, for ease of exposition, we will assume that no
arguments are passed on the stack.
%% \rb{Conceivably on alloc
%%   operations, etc., which we have not yet mentioned.}

@@

\item From callee back to caller, when execution is about to reach a
  return point.  The exit sequence retags all locations in the
  function frame with $\tagNoDepth$. At the actual return instruction,
  the $\PCname$ tag is decremented. \leo{Not sure the actual policy
    does that now (retag unused). But I don't think it's a
    confidentiality or integrity violation.}
  \rb{Correct, and it doesn't. Some versions of the policy do clear
    the stack, retagging would be easily added.}

@@

We validated the micro-policy of the previous section by systematically
testing that it satisfies eager stack integrity and
confidentiality. We used a Coq specification of the RISC-V
architecture \leo{cite: https://github.com/mit-plv/riscv-coq} and
extended it with a runtime monitor implementing the stack-safety
micro-policy.\leo{Footnote: code will be made public}\bcp{For the reviewers?
 Or later?} We chose the
Coq proof assistant as the setting for our implementation for two
reasons: first, to ensure that our coinductive trace definitions are
well-formed, and reason about them\leo{Do we prove anything
  interesting anymore?}, and to leverage the power of the QuickChick
property-based testing framework~\citep{Pierce:SF4}.

@@

To write such generators we build on the work of \leo{ICFP 2013},
where they introduced {\em generation by execution} to produce
interesting programs. Generation by execution receives as an input a
partially instantiated machine state and attempts to generate an
instruction (or a sequence of instructions such as a blessed
sequence) that makes sense locally (e.g., jumps go to a potentially valid
code location and loads read from from a potentially valid stack
location). Then
we step the machine and repeat the process until we generate or
execute some target number of instructions, or if we reach a point
where the machine can't step anymore.

@@

@@

To evaluate the effectiveness of testing against our formal properties, we
used {\em mutation testing}\leo{cite}. In mutation testing, we inject errors
(mutations) in a program that should cause the property of interest (here,
stack safety) to fail, and ensure that the testing framework can find
them. The bugs we used for our evaluation were either artificially generated
by us (deliberately weakening the micro-policy in ways that we expected
should break its guarantees), or actual bugs that we discovered through
testing our implementation. We elaborate on some such bugs below.

The mean-time-to-failure (MTTF) for various bugs can be found in \leo{Table
  with numbers?}\bcp{Sure.}. All experiments were run in a \leo{Machine
  spec}. \leo{Something like: There is really no baseline to compare this
  against, but include it here for the future/reasons?}
